---
title: "Current Economic Outlook"
date: "`r paste0('updated: ',as.Date(file.info(here::here('processed_data','nested_dataframe.rds'))$mtime-lubridate::hours(7)))`"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    social: menu
    source_code: https://github.com/bcgov/current_economic_analysis
    css: style.css
runtime: shiny
resource_files:
- processed_data/nested_dataframe.rds
- R/functions.R
- processed_data/for_heatmap.rds
- processed_data/for_up_down.rds
- processed_data/commentary.rds
---

```{r global, include=FALSE}
# Copyright 2022 Province of British Columbia
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
# http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and limitations under the License.
# libraries----------
library(tidyverse)
library(tsibble) # data includes a tsibble::yearquarter field
library(ggthemes)
library(markdown)
library(formattable)
library(ggplot2)
library(conflicted)
library(plotly)
library(heatmaply)
conflicts_prefer(dplyr::filter)
conflicts_prefer(formattable::style)
theme_set(theme_minimal(base_size = 15, base_family = "Franklin Gothic Book"))

# load objects----------
commentary <- readRDS(here::here("processed_data", "commentary.rds"))
nested_dataframe <- readRDS(here::here("processed_data", "nested_dataframe.rds"))
for_heatmap <- readRDS(here::here("processed_data", "for_heatmap.rds"))
for_up_down <- readRDS(here::here("processed_data", "for_up_down.rds"))
source(here::here("R", "functions.R"))

```

# Commentary:

## Inputs {.sidebar}

<br><br>

-   This dashboard retrieves time series pertinent to British Columbia's current economic outlook from `r dim(nested_dataframe)[1]` sources.
-   The data is updated nightly (with the exception of building permits: weekly)

<br>

##

###

<br><br>

**New Releases:**

```{r, results='asis'}
fresh <- commentary%>%
            filter(relatively=="fresh")
if(nrow(fresh)==0){
  cat("* No new data over last 24 hours.")
}else{
  cat(paste("*", 
            fresh%>%
              pull()),
            sep = "\n")
}  
```

**Notably High:**

```{r, results='asis'}
cat(paste("*", 
          commentary%>%
            filter(relatively=="high")%>%
            pull()), 
    sep = "\n")
```

**Notably Low:**

```{r, results='asis'}
cat(paste("*", 
          commentary%>%
            filter(relatively=="low")%>%
            pull()),
    sep = "\n")
```

# Most recent changes:

## Inputs {.sidebar}

<br><br>


```{r}
sliderInput("change_thres", "Changes smaller than this in grey.",
  min = 0, max = 100, value = 5, post = "%"
)
sliderInput("annual_change_thres", "Annual changes smaller than this in grey.",
  min = 0, max = 100, value = 50, post = "%"
)

```

## Row

```{r}
change_thres <- .1
annual_change_thres <- .1
renderFormattable({
  formattable(for_up_down, 
              align=c("l","r","r","c","c"),
              list(
    Value = formatter("span", x ~ case_when(str_detect(for_up_down$Name,"Merchandise Trade")~ scales::dollar(x, accuracy=1),
                                            str_detect(for_up_down$Name,"dollars")~ scales::dollar(x, accuracy=1),
                                            str_detect(for_up_down$Name,"yield")~ scales::percent(x, accuracy=.01),
                                            str_detect(for_up_down$Name,"rate")~ scales::percent(x, accuracy=.01),
                                            str_detect(for_up_down$Name,"Exchange Rate")~ scales::dollar(x, accuracy=.01),
                                            TRUE ~ scales::comma(x, accuracy = 1))
                      ),                  
   Change = formatter("span",
      style = x ~ style(color = case_when(
        x + input$change_thres / 100 < 0 ~ "red",
        x - input$change_thres / 100 > 0 ~ "green",
        TRUE ~ "gainsboro"
      )),
      x ~ icontext(case_when(
        x + input$change_thres / 100 < 0 ~ "arrow-down",
        x - input$change_thres / 100 > 0 ~ "arrow-up",
        TRUE ~ "arrow-right"
      ))
    ),
    `Annual Change` = formatter(
      "span",
      style = x ~ style(color = case_when(
        x + input$annual_change_thres / 100 < 0 ~ "red",
        x - input$annual_change_thres / 100 > 0 ~ "green",
        TRUE ~ "gainsboro"
      )),
      x ~ icontext(case_when(
        x + input$annual_change_thres / 100 < 0 ~ "arrow-down",
        x - input$annual_change_thres / 100 > 0 ~ "arrow-up",
        TRUE ~ "arrow-right"
      ))
    )
  ))
})
```

# Individual Time Series:

## Inputs {.sidebar}
<br><br>

```{r}
selectInput("name",
  label = "Choose a data set:",
  choices = sort(nested_dataframe$name, decreasing = TRUE),
  selected = sort(nested_dataframe$name, decreasing = TRUE)[1]
)
selectInput(
  "type",
  "Choose a plot type:",
  c("line", "column", "smooth"),
  selected = "line",
  multiple = FALSE
)
conditionalPanel(
  condition = "input.type == 'smooth'",
  sliderInput(
    "span",
    "how smooth?",
    .1,
    .9,
    .5
  )
)
conditionalPanel(
  condition = "input.type == 'smooth'",
  sliderInput(
    "alpha",
    "Alpha for raw data?",
    .1,
    .9,
    .25
  )
)

selectInput(
  "pal",
  "Choose a colour palette:",
  c("Viridis", "Set2", "Paired", "Dark2"),
  selected = "Dark2",
  multiple = FALSE
)

checkboxInput("include_origin", "Should plot include origin?", FALSE)
checkboxInput("facet", "Do you want to facet by Series?", FALSE)
reactive_df <- reactive({
  extract_cell(nested_dataframe, input$name)
})
```

## Row {data-height=925 .tabset}
-------------------------------------

### Level:

```{r}
plotly::renderPlotly({ 
  plotly_ts(
    reactive_df(),
    Value,
    "unknown",
    "the level of ",
    input$type,
    input$pal,
    title = input$name,
    spn = input$span,
    el = input$include_origin,
    alf = input$alpha,
    facet =input$facet
  )
}) %>%
  bindCache(
    "level",
    input$name,
    input$type,
    input$pal,
    input$span,
    input$include_origin,
    input$alpha,
    input$facet
  )
```

### Change from last observation:

```{r}
plotly::renderPlotly({
  plotly_ts(
    reactive_df(),
    `Change`,
    "percent",
    "the rate of change of ",
    input$type,
    input$pal,
    title = input$name,
    spn = input$span,
    alf =input$alpha,
    facet = input$facet
  )
}) %>%
  bindCache(
    "change",
    input$name,
    input$type,
    input$pal,
    input$span,
    input$alpha,
    input$facet
  )
```

### Annual Change:

```{r}
plotly::renderPlotly({
  plotly_ts(
    reactive_df(),
    `Annual Change`,
    "percent",
    "the annual rate of change of ",
    input$type,
    input$pal,
    title = input$name,
    spn = input$span,
    alf =input$alpha,
    facet =input$facet
  )
}) %>%
  bindCache(
    "annual change",
    input$name,
    input$type,
    input$pal,
    input$span,
    input$alpha,
    input$facet
  )
```

### The Data:

```{r}
DT::renderDataTable(server = FALSE, {
  reactive_df() %>%
    select(-Source) %>%
    mutate(
      `Change` = scales::percent(`Change`, accuracy = .1),
      `Annual Change` = scales::percent(`Annual Change`, accuracy = .1)
    ) %>%
    DT::datatable(
      extensions = "Buttons",
      rownames = FALSE,
      options = list(
        columnDefs = list(list(className = "dt-center", targets = "_all")),
        paging = TRUE,
        scrollX = TRUE,
        scrollY = TRUE,
        searching = TRUE,
        ordering = TRUE,
        dom = "Btip",
        buttons = list(
          list(extend = "csv", filename = input$name),
          list(extend = "excel", filename = input$name)
        ),
        pageLength = 10,
        lengthMenu = c(3, 5)
      )
    )
}) %>%
  bindCache(input$name)
```

Row {data-height=75}
-------------------------------------

###

```{r}
renderUI({
  if (str_detect(reactive_df()$Source[1], "http")) {
    HTML(paste("Source: <a href='", reactive_df()$Source[1], "'>", reactive_df()$Source[1], "</a>"))
  } else {
    paste("Source: ", reactive_df()$Source[1])
  }
})
```

# All monthly series:

## Inputs {.sidebar data-width="325"}

<br>

-   The levels of the time series differ by several orders of magnitude.
-   Levels can also be close to zero implying that growth rates can also vary by several orders of magnitude.
-   Two data normalizations are considered:
    -   Binning: within-series percentile scores.
    -   Re-scaling: via transformation $(x-\min(x))/(\max(x)-\min(x))$
-   Note that re-scaling retains information regarding magnitude, which binning discards.

```{r}
selectInput("norm",
  label = "Choose a data visualization:",
  choices = c(
    "Binned Level",
    "Rescaled Level",
    "Binned Change",
    "Rescaled Change",
    "Binned Annual Change",
    "Rescaled Annual Change"
  ),
  selected = "Rescaled Value"
)
```

## Column

### 

```{r}
renderPlotly({
  wide <- for_heatmap %>%
    filter(`Period Starting`> today() - years(10))%>%
    select(longname, `Period Starting`, input$norm) %>%
    pivot_wider(id_cols = longname, names_from = "Period Starting", values_from = input$norm) %>%
    column_to_rownames(var = "longname")
 
  heatmaply::heatmaply(wide, fontsize_row = 6, fontsize_col = 4, dendrogram="row")

})%>%
  bindCache(
    input$norm
  )
```


