---
title: "Current Economic Outlook"
date: "`r paste0('updated: ',as.Date(file.info(here::here('processed_data','nested_dataframe.rds'))$mtime-lubridate::hours(7)))`"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    social: menu
    source_code: https://github.com/bcgov/current_economic_analysis
    css: style.css
runtime: shiny
resource_files:
- processed_data/mpi.rds
- processed_data/nested_dataframe.rds
- R/functions.R
- processed_data/for_heatmap.rds
- processed_data/for_up_down.rds
- processed_data/commentary.rds
---

```{r global, include=FALSE}
# Copyright 2022 Province of British Columbia
#
# Licensed under the Apache License, Version 2.0 (the "License");
# you may not use this file except in compliance with the License.
# You may obtain a copy of the License at
#
# http://www.apache.org/licenses/LICENSE-2.0
#
# Unless required by applicable law or agreed to in writing, software distributed under the License is distributed on an "AS IS" BASIS,
# WITHOUT WARRANTIES OR CONDITIONS OF ANY KIND, either express or implied.
# See the License for the specific language governing permissions and limitations under the License.
# libraries----------
# library(DT)
library(tidyverse)
library(tsibble) # data includes a tsibble::yearquarter field
#library(lubridate) # years and months not exported by lubridate
library(ggthemes)
library(markdown)
library(formattable)
library(ggplot2)
#library(Formula)
#library(inum)
#library(libcoin)
#library(partykit)

theme_set(theme_minimal(base_size = 15, base_family = "Franklin Gothic Book"))

# load objects----------
commentary <- readRDS(here::here("processed_data", "commentary.rds"))
nested_dataframe <- readRDS(here::here("processed_data", "nested_dataframe.rds"))
for_heatmap <- readRDS(here::here("processed_data", "for_heatmap.rds"))
for_up_down <- readRDS(here::here("processed_data", "for_up_down.rds"))
mpi <- readRDS(here::here("processed_data", "mpi.rds")) %>%
  filter(last_update > today() - years(5)) # a single project had a last_update value of 2016?
source(here::here("R", "functions.R"))
```

# Commentary:

## Inputs {.sidebar}

<br><br>

-   This dashboard retrieves time series pertinent to British Columbia's current economic outlook from `r dim(nested_dataframe)[1]` sources.
-   The data is updated nightly (with the exception of building permits: weekly)

<br>

## Column

<br><br>

**New Releases:**

```{r, results='asis'}
fresh <- commentary%>%
            filter(relatively=="fresh")
if(nrow(fresh)==0){
  cat("* No new data over last 24 hours.")
}else{
  cat(paste("*", 
            fresh%>%
              pull()),
            sep = "\n")
}  
```

**Notably High:**

```{r, results='asis'}
cat(paste("*", 
          commentary%>%
            filter(relatively=="high")%>%
            pull()), 
    sep = "\n")
```

**Notably Low:**

```{r, results='asis'}
cat(paste("*", 
          commentary%>%
            filter(relatively=="low")%>%
            pull()),
    sep = "\n")
```

# Most recent changes:

## Inputs {.sidebar}

<br><br>


```{r}
sliderInput("change_thres", "Changes smaller than this in grey.",
  min = 0, max = 100, value = 5, post = "%"
)
sliderInput("annual_change_thres", "Annual changes smaller than this in grey.",
  min = 0, max = 100, value = 50, post = "%"
)

```

## Column

```{r}
change_thres <- .1
annual_change_thres <- .1
renderFormattable({
  formattable(for_up_down, 
              align=c("l","r","r","c","c"),
              list(
    Value = formatter("span", x ~ case_when(str_detect(for_up_down$Name,"Merchandise Trade")~ scales::dollar(x, accuracy=1),
                                            str_detect(for_up_down$Name,"dollars")~ scales::dollar(x, accuracy=1),
                                            str_detect(for_up_down$Name,"yield")~ scales::percent(x, accuracy=.01),
                                            str_detect(for_up_down$Name,"rate")~ scales::percent(x, accuracy=.01),
                                            str_detect(for_up_down$Name,"Exchange Rate")~ scales::dollar(x, accuracy=.01),
                                            TRUE ~ scales::comma(x, accuracy = 1))
                      ),                  
   Change = formatter("span",
      style = x ~ style(color = case_when(
        x + input$change_thres / 100 < 0 ~ "red",
        x - input$change_thres / 100 > 0 ~ "green",
        TRUE ~ "gainsboro"
      )),
      x ~ icontext(case_when(
        x + input$change_thres / 100 < 0 ~ "arrow-down",
        x - input$change_thres / 100 > 0 ~ "arrow-up",
        TRUE ~ "arrow-right"
      ))
    ),
    `Annual Change` = formatter(
      "span",
      style = x ~ style(color = case_when(
        x + input$annual_change_thres / 100 < 0 ~ "red",
        x - input$annual_change_thres / 100 > 0 ~ "green",
        TRUE ~ "gainsboro"
      )),
      x ~ icontext(case_when(
        x + input$annual_change_thres / 100 < 0 ~ "arrow-down",
        x - input$annual_change_thres / 100 > 0 ~ "arrow-up",
        TRUE ~ "arrow-right"
      ))
    )
  ))
})
```

# Individual Time Series:

## Inputs {.sidebar}
<br><br>

```{r}
selectInput("name",
  label = "Choose a data set:",
  choices = sort(nested_dataframe$name, decreasing = TRUE),
  selected = sort(nested_dataframe$name, decreasing = TRUE)[1]
)
selectInput(
  "type",
  "Choose a plot type:",
  c("line", "column", "smooth"),
  selected = "line",
  multiple = FALSE
)
conditionalPanel(
  condition = "input.type == 'smooth'",
  sliderInput(
    "span",
    "how smooth?",
    .1,
    .9,
    .5
  )
)
conditionalPanel(
  condition = "input.type == 'smooth'",
  sliderInput(
    "alpha",
    "Alpha for raw data?",
    .1,
    .9,
    .25
  )
)

selectInput(
  "pal",
  "Choose a colour palette:",
  c("Viridis", "Set2", "Paired", "Dark2"),
  selected = "Dark2",
  multiple = FALSE
)

checkboxInput("include_origin", "Should plot include origin?", FALSE)
checkboxInput("facet", "Do you want to facet by Series?", FALSE)
reactive_df <- reactive({
  extract_cell(nested_dataframe, input$name)
})
```

## Column {.tabset}

### Level:

#### 

```{r}
plotly::renderPlotly({ 
  plotly_ts(
    reactive_df(),
    Value,
    "unknown",
    "the level of ",
    input$type,
    input$pal,
    title = input$name,
    spn = input$span,
    el = input$include_origin,
    alf = input$alpha,
    facet =input$facet
  )
}) %>%
  bindCache(
    "level",
    input$name,
    input$type,
    input$pal,
    input$span,
    input$include_origin,
    input$alpha,
    input$facet
  )
renderUI({
  if (str_detect(reactive_df()$Source[1], "http")) {
    HTML(paste("Source: <a href='", reactive_df()$Source[1], "'>", reactive_df()$Source[1], "</a>"))
  } else {
    paste("Source: ", reactive_df()$Source[1])
  }
})
```

### Change from last observation:

#### 

```{r}
plotly::renderPlotly({
  plotly_ts(
    reactive_df(),
    `Change`,
    "percent",
    "the rate of change of ",
    input$type,
    input$pal,
    title = input$name,
    spn = input$span,
    alf =input$alpha,
    facet = input$facet
  )
}) %>%
  bindCache(
    "change",
    input$name,
    input$type,
    input$pal,
    input$span,
    input$alpha,
    input$facet
  )
renderUI({
  if (str_detect(reactive_df()$Source[1], "http")) {
    HTML(paste("Source: <a href='", reactive_df()$Source[1], "'>", reactive_df()$Source[1], "</a>"))
  } else {
    paste("Source: ", reactive_df()$Source[1])
  }
})
```

### Annual Change:

#### 

```{r}
plotly::renderPlotly({
  plotly_ts(
    reactive_df(),
    `Annual Change`,
    "percent",
    "the annual rate of change of ",
    input$type,
    input$pal,
    title = input$name,
    spn = input$span,
    alf =input$alpha,
    facet =input$facet
  )
}) %>%
  bindCache(
    "annual change",
    input$name,
    input$type,
    input$pal,
    input$span,
    input$alpha,
    input$facet
  )
renderUI({
  if (str_detect(reactive_df()$Source[1], "http")) {
    HTML(paste("Source: <a href='", reactive_df()$Source[1], "'>", reactive_df()$Source[1], "</a>"))
  } else {
    paste("Source: ", reactive_df()$Source[1])
  }
})
```

### The Data:

#### 

```{r}
DT::renderDataTable(server = FALSE, {
  reactive_df() %>%
    select(-Source) %>%
    mutate(
      `Change` = scales::percent(`Change`, accuracy = .1),
      `Annual Change` = scales::percent(`Annual Change`, accuracy = .1)
    ) %>%
    DT::datatable(
      extensions = "Buttons",
      rownames = FALSE,
      options = list(
        columnDefs = list(list(className = "dt-center", targets = "_all")),
        paging = TRUE,
        scrollX = TRUE,
        scrollY = TRUE,
        searching = TRUE,
        ordering = TRUE,
        dom = "Btip",
        buttons = list(
          list(extend = "csv", filename = input$name),
          list(extend = "excel", filename = input$name)
        ),
        pageLength = 10,
        lengthMenu = c(3, 5)
      )
    )
}) %>%
  bindCache(input$name)
renderUI({
  if (str_detect(reactive_df()$Source[1], "http")) {
    HTML(paste("Source: <a href='", reactive_df()$Source[1], "'>", reactive_df()$Source[1], "</a>"))
  } else {
    paste("Source: ", reactive_df()$Source[1])
  }
})
```

# All monthly series:

## Inputs {.sidebar data-width="325"}

<br>

-   The levels of the time series differ by several orders of magnitude.
-   Levels can also be close to zero implying that growth rates can also vary by several orders of magnitude.
-   Two data normalizations are considered:
    -   Binning: within-series percentile scores.
    -   Re-scaling: via transformation $(x-\min(x))/(\max(x)-\min(x))$
-   Note that re-scaling retains information regarding magnitude, which binning discards.

```{r}
selectInput("norm",
  label = "Choose a data visualization:",
  choices = c(
    "Binned Level",
    "Rescaled Level",
    "Binned Change",
    "Rescaled Change",
    "Binned Annual Change",
    "Rescaled Annual Change"
  ),
  selected = "Rescaled Value"
)
```

## Column

### 

```{r}
renderPlot({
  wide <- for_heatmap %>%
    filter(`Period Starting`> today() - years(10))%>%
    select(longname, `Period Starting`, input$norm) %>%
    pivot_wider(id_cols = longname, names_from = "Period Starting", values_from = input$norm) %>%
    column_to_rownames(var = "longname")
 
  pheatmap::pheatmap(wide,
    cluster_cols = FALSE,
    angle_col = 0,
    main = str_to_title(paste("Monthly data, ", input$norm)),
    col = viridisLite::viridis(100),
    labels_col = if_else(lubridate::month(lubridate::ymd(colnames(wide))) == 1,
      as.character(lubridate::year(lubridate::ymd(colnames(wide)))),
      ""))
})%>%
  bindCache(
    input$norm
  )
```

# MPI:

## Inputs {.sidebar}

```{r}
char_cols <- colnames(mpi)[sapply(mpi, class) == "character"]
possible_vars <- char_cols[char_cols != "source"]

selectInput("x_var",
  label = "Choose a variable for x axis:",
  choices = possible_vars, selected = "region"
)
selectInput("y_var",
  label = "Choose a variable for y axis:",
  choices = possible_vars, selected = "project_stage"
)
selectInput("facet_var",
  label = "Choose a variable to facet by:",
  choices = possible_vars, selected = "project_status"
)

observe({
  updateSelectInput(session, "y_var", choices = possible_vars[possible_vars != input$x_var])
})
observe({
  updateSelectInput(session, "facet_var", choices = possible_vars[possible_vars != input$x_var & possible_vars != input$y_var])
})

mpi_group_summary <- reactive({
  mpi %>%
    group_by_(
      `X-axis variable` = input$x_var, # group_by_ DEPRECATED.
      `Y-axis variable` = input$y_var,
      `Facet variable` = input$facet_var
    ) %>%
    summarize(`Total Cost (M)` = sum(estimated_cost, na.rm = TRUE)) %>%
    filter(`Total Cost (M)` > 0)
})
```

## Column {.tabset}

### Plot:

```{r}
renderPlot({
  plt <- ggplot(mpi_group_summary(), aes(`X-axis variable`,
    `Y-axis variable`,
    fill = `Total Cost (M)`,
    text = paste("Total project cost is", scales::comma(`Total Cost (M)`), "Million.")
  )) +
    geom_tile() +
    scale_fill_viridis_c(trans = "log10", labels = scales::comma) +
    facet_wrap(~`Facet variable`) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) +
    labs(
      x = input$x_var,
      y = input$y_var,
      title = paste("Major Project Inventory:", tsibble::yearmonth(mpi$last_update[1]))
    )
  aest_fix_labs(plt)
}) %>%
  bindCache(input$filter_var, input$x_var, input$y_var, input$facet_var)
```

### Table:

```{r}
DT::renderDataTable(server = FALSE, {
  mpi_group_summary() %>%
    DT::datatable(
      extensions = "Buttons",
      options = list(
        rownames = FALSE,
        columnDefs = list(list(className = "dt-center", targets = "_all")),
        paging = TRUE,
        scrollX = TRUE,
        scrollY = TRUE,
        searching = TRUE,
        ordering = TRUE,
        dom = "Btip",
        buttons = list(
          list(extend = "csv", filename = "MPI"),
          list(extend = "excel", filename = "MPI")
        ),
        pageLength = 10,
        lengthMenu = c(3, 5)
      )
    )
}) %>%
  bindCache(input$filter_var, input$x_var, input$y_var, input$facet_var)
```
